What does the crowd believe? A hierarchical approach to estimating subjective beliefs from empirical data

### 2016-01-31 text/01_CogSci_abstract/paper.tex
add watch example bins
intro rewrite

\definecolor{Green}{RGB}{10,200,100}
\newcommand{\ndg}[1]{\textcolor{Green}{[ndg: #1]}}  
\definecolor{Blue}{RGB}{10,100,200}
\newcommand{\jd}[1]{\textcolor{Blue}{[jd: #1]}}

@@

\todo[inline]{Each author's name should appear
on a separate line, 11~point bold, and centered, with the author's
email address in parentheses. Under each author's name list the
author's affiliation and postal address in ordinary 10~point type.}

\todo[inline]{who is an author? which order?}

\begin{abstract}
  Experimental participants' beliefs about everyday events are of theoretical interest on their own and also an
  important ingredient for model building---especially in Bayesian models of phenomena such as logical
  reasoning, future predictions, and language use. Here, we explore several recently used 
  methods for measuring subjective beliefs about unidimensional properties. \ndg{say what kind of subjective beliefs /distributions we look at more clearly.} 
  We a hierarchical Bayesian data analysis model for inferring likely ``population-level beliefs'' as the central
  tendency of participants' individual-level beliefs. 
  Three different dependent measures---\ndg{name them?}---are related to individual-level beliefs through likelihood ``linking functions.''
%  functions for three types of task measures. 
  Our results suggest that the practice of
  using averaged normalized slider ratings for binned quantities is a practical and fairly good
  approximator of inferred population-level beliefs.
  Other dependent measures yield less reliable results.
\textbf{Keywords:} subjective beliefs, hierarchical modeling, Bayesian data analysis, Bayesian
cognitive models 
\end{abstract}

@@

We cannot look into a person's head. But in making sense of observed behavior we readily
ascribe beliefs and desires to fellow agents. This happens intuitively, in folk psychology, but
also in science. Scientific ascription of latent mental states plays an important role in many
explanations of higher-order cognition: decision making, planning, reasoning,
and language use. Naturally, then, it becomes important to validate any method for eliciting mental states, 
to be used for explanatory purposes.

A family of models where this is particularly pressing are Bayesian models of cognition which
seek to explain task behavior in a variety of domains as partially informed by what participants
believe about mundane events---their \emph{prior beliefs}. Take interpretation of language. Empirical data on whether a
statement like ``That watch cost $n$ dollars'' is understood to convey speaker affect, rather
than literal meaning, can be explained well by a Bayesian model of utterance interpretation
\citep{KaoWu2014:Nonliteral-Unde}. In this model a crucial role is played by
an empirical measure of participants' expectations about the likely or normal price of a
watch. 
Other examples of domains in which empirically successful models have included some
measure of participants' belief include making predictions about everyday events 
\citep{GriffithsTenenbaum2006:Optimal-Predict}, the strength of pragmatic enrichments
\citep{DegenTessler2015:Wonky-worlds:-L}, the interpretation of vague quantifiers
\citep{SchollerFranke2015:Semantic-values}, \textcolor{red}{ \dots }

\todo[inline]{fill me}

Many methods of assessing subjective beliefs exist. One approach is to take actual frequencies
as an approximation \citep{GriffithsTenenbaum2006:Optimal-Predict}. This will not
work when sufficient real-world data is unavailable---for instance, for beliefs about imaginary events. 
Even where available, real-world
frequencies may deviate systematically from participants' beliefs, and this can be problematic for
the predictions of a model that relies on subjective beliefs.
%\citep{MarcusDavis2013:How-Robust-Are-}.

Another approach is to experimentally elicit relevant beliefs. Many techniques for this
exist, especially in the economics literature
\citep{MorganHenrion1990:Uncertainty:-A-,Manski2004:Measuring-Expec,SchlagTremewan2014:A-penny-for-you,AndersenFountain2014:Estimating-Subj}. The
perhaps most prominent method uses so-called \emph{scoring rules}
\citep{Savage1971:Elicitation-of-,SchlagTremewan2014:A-penny-for-you}. Roughly speaking,
scoring rules are particular schemes of rewarding participants' answers in such a way as to
make sure that responses are what the experimenter wants them to be: the actual subjective probability of an
event, the mean of a distribution, a credible interval etc. Scoring rules are highly faithful
but require participants to become sufficiently familiar with the (usually probabilistic)
payoff scheme. Scoring rules also must make assumptions about how much participants value
certain probabilistic gambles and may therefore require elaborate means of data analysis
\citep{AndersenFountain2014:Estimating-Subj}. Similar practical concerns apply to other methods
as well, such as the use of \emph{iterated learning} for prior estimation
\citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.

For common applications, many of these procedures may be excessively complex or too demanding
on both researcher and participant. It would be beneficial if there was an easily applicable
technique of eliciting subjective beliefs that is good enough for testing the general
predictions of models that rely on these measures for their predictions. A further complication
is that it may be impractical to obtain, from the same participant, data on subjective beliefs
and data on the task of interest, especially if exposure to one of these measures is likely to
affect the other. This suggests that, for practical reasons, there is a demand for
reliable-enough measures of what we will  call ``population-level beliefs'': estimates
of the central tendency or average of subjective beliefs in a given population.

A very simple technique that has been used with apparent success is to have participants adjust
sliders in order to express their (relative) levels of support for values of uncertain
contiguous quantities
\citep{KaoWu2014:Nonliteral-Unde,SchollerFranke2015:Semantic-values}. \textcolor{red}{ MORE REFERENCES } Participants are confronted with a short and simple description of a situation or an
event, such as ``John bought a watch,'' and asked for their estimates of the watch's price. To
give their estimate, they are asked to adjust sliders whose endpoints are labelled as ``extremely
likely'' and ``extremely unlikely,'' for a number of bins that partition the range of plausible
values (established by a pre-test). E.g., for the price of a watch bins could be \jd{erin: fill in}
\textcolor{red}{ FILL ME }

@@

\ndg{say more clearly that we focus on unidimensional distributions and why.}

@@

>>>>

\definecolor{Green}{RGB}{10,200,100}
\newcommand{\ndg}[1]{\textcolor{Green}{[ndg: #1]}}  
\definecolor{Blue}{RGB}{10,100,200}
\newcommand{\jd}[1]{\textcolor{Blue}{[jd: #1]}} 
\definecolor{Red}{RGB}{178,34,34}
\newcommand{\mf}[1]{\textcolor{Red}{[mf: #1]}}

@@

\begin{abstract}
  Experimental participants' beliefs about everyday events are of theoretical interest on their
  own and also an important ingredient for model building---especially in Bayesian models of
  phenomena such as logical reasoning, future predictions, and language use. Here, we explore
  several recently used methods for measuring subjective beliefs about unidimensional
  contiguous properties, such as the likely price of a new watch. We use a hierarchical
  Bayesian model for inferring likely ``population-level beliefs'' as the central tendency of
  participants' individual-level beliefs.  Three different dependent measures are related to
  individual-level beliefs through likelihood ``linking functions:'' (i) judgements of
  (relative) likelihood of intervals of values via continuous sliders, (ii) a give-a-number
  task, and (iii) choice of the more likely of two intervals of values. Our results suggest
  that using averaged normalized slider ratings for binned quantities is a practical and fairly
  good approximator of inferred population-level beliefs.

  \textbf{Keywords:} subjective beliefs, hierarchical modeling, Bayesian data analysis,
  Bayesian cognitive models
\end{abstract}

@@

We cannot look into a person's head. But in making sense of observed behavior we readily
ascribe beliefs and desires to fellow agents. This happens intuitively, in folk psychology, but
also in science. Scientific ascription of latent mental states plays an important role in many
explanations of higher-order cognition: decision making, planning, reasoning, and language
use. It is therefore vital to have methods for validating explanatory mental states
ascriptions.

A family of models where this is particularly pressing are Bayesian models of cognition which
seek to explain task behavior in a variety of domains as partially informed by what
participants believe about mundane events---their \emph{prior beliefs}. Take interpretation of
language. Empirical data on whether a statement like ``That watch cost $n$ dollars'' is
understood to convey speaker affect, rather than literal meaning, can be explained well by a
Bayesian model of utterance interpretation \citep{KaoWu2014:Nonliteral-Unde}. In this model a
crucial role is played by an empirical measure of participants' expectations about the likely
or normal price of a watch.  Other examples of domains in which empirically successful models
have included a measure of participants' prior beliefs include making predictions about
everyday events \citep{GriffithsTenenbaum2006:Optimal-Predict}, referential communication
\citep{FrankGoodman2012:Predicting-Prag}, the strength of pragmatic enrichments
\citep{DegenTessler2015:Wonky-worlds:-L}, or the interpretation of vague quantifiers
\citep{SchollerFranke2015:Semantic-values}. \mf{anyone wants any other references here?}

Many methods of assessing subjective beliefs exist. One approach is to take actual frequencies
as an approximation to subjective beliefs
\citep{GriffithsTenenbaum2006:Optimal-Predict}. Unfortunately, the relevant data may not be
available (e.g., for one-shot or imaginary events) and real-world frequencies may deviate from
participants' subjective beliefs for many reasons.

Another possibility is to empirically measure subjective beliefs by ``give-a-number''
tasks. The simplest version would be this: being told that John just bought a new watch,
subjects are asked for a single numerical estimate of its price. This task is easy to
comprehend and implement, but a single number does not provide much information about the
subjective belief that it is (arguably) a manifestation of. One solution is to infer which
parameterized distribution best explains the observed number choices, either at the individual
level \citep{Manski2004:Measuring-Expec} or on the population level
\citep{TauberSteyvers2013:Inferring-Subje}.

More sophisticated ``give-a-number'' tasks give more information, but can be rather complex and
more difficult to implement. A prominent method in economics uses \emph{scoring rules}
\citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}. Roughly
put, scoring rules are schemes of rewarding participants' answers that make sure that responses
(by rational subjects) faithfully reflect a measure of interest: the actual subjective
probability of an event, the mean of a distribution, a credible interval etc. Scoring-rule
tasks require participants to become sufficiently familiar with the (usually probabilistic)
payoff scheme. Another insightful, but also complex approach is to use a cleverly designed
\emph{iterated learning paradigm} \citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.

In sum, there is tradeoff between simplicity of paradigms and their informativity. Ideally, we
would like to have an experimental method for measuring subjective beliefs that (i) gives us
sufficient and reliable-enough information about subjective beliefs to derive testable
predictions from cognitive models that rely on such information, (ii) is easy to understand by
the participants, (iii) is easy to implement by researchers whose focus and expertise is not
belief-elicitation as such, and that (iv) does not rely on sophisticated methods of extracting
the relevant information from the observed data. Moreover, the ideal method would be (v)
flexible enough to allow inferred subjective belief distributions beyond standard parameterized
distributions and (vi) would provide us with a reasonable approximation of a ``population-level
belief,'' i.e., an estimate of the central tendency or average of subjective beliefs in a given
population. The latter allows to use one set of subjects for measuring prior beliefs and
another for whatever other task is of interest.

A simple technique that seems to fit this bill is the \emph{binned histograms} task that has
recently been used with apparent success
\citep{KaoWu2014:Nonliteral-Unde,SchollerFranke2015:Semantic-values}. \mf{someone please add
  references: where did you use binned sliders?} Participants adjust sliders in order to
express their (relative) subjective beliefs about how likely it is that the value of an
uncertain contiguous quantity lies in some interval of possible values (a bin). E.g.,
participants read that John bought a watch.  To report their beliefs about the watch's price,
they adjust sliders whose endpoints are labelled as ``extremely likely'' and ``extremely
unlikely.'' Each slider corresponds to on out of 15 bins that partition the range of plausible
values (established by a pre-test), such as 0--50\$, 50--100\$, \dots 700--750\$. Each
subject's slider ratings are normalized within each item, and the results are averaged across
subjects to obtain an estimate of the population's average belief about each item.

@@

@@

\begin{itemize}
\item \ndg{say more clearly that we focus on unidimensional distributions and why.}
\end{itemize}
